DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

\mdeff{measure the computational time of filtering on a 1/12, 1/48 or 1/192 of the sphere with the graph. Add the result to Figure 4 and compare with filtering of the full sphere (with SHT and graph).}

\subsection{}

\begin{mdframed}[style=comment]
Secondly, using only SVM as a comparison might skew how well this technique appears to works in a deep learning setting. In principle (and I say this not knowing for sure) 2D CNNs should probably be able to work as well as DeepSphere (although no where near as elegantly) by taking in a projected 2D image of the sky and using *loads* of filters to learn about the distortion at different parts of the projected sky. It would be interesting (and I think a fairly easy test) to check how well DeepSphere performs against a 2D CNN on projections.
\end{mdframed}

\todo{Assigned: @Nati, @Michael, @Tomek}
\nati{Please check if you like what I did in light of the discussion we had.}

We share the opinion of the reviewer when he says that 2D CNNs should probably work as well as DeepSphere, ``by taking in a projected 2D image of the sky and using *loads* of filters to learn about the distortion at different parts of the projected sky.'' Hence we decided to add a new experiment in the paper.

Nevertheless, making a comparison is technically not as trivial as it may seems. 
Problematically a radial projection is computationally expensive to the point where, because of the dataset augmentation, it becomes significantly more expensive than the actual training of the network.
Hence, we decided to use the geometrical properties of HEALPix to create a ``special'' projection. 
In the HEALPix sampling, the sphere is first separated into $12$ parts that are then subdivided into $4$ recursively.
Eventually, each of these $12$ part resembles a square grid. 
As the problem we are solving only involve one of these 12 parts, we can use this natural structure to simply re-assign each HEALPix pixel to one of a 2 dimensional grid of size $N_{sides} \times N_{sides}$. 
This operation is computationally very cheap and probably spread quite evenly across the resulting image. 
This projection has recently been used for a CNN in cosmology in the very recent work \cite{krachmalnicoff2019convolutional}.

As we believe that this projection is the optimal one to use a traditonal CNN for application, we performed the experiment using it.

Our expectation were similar with the one of the reviewer, but the results did surprise us as the performance of the traditional CNN did in general not match DeepSphere, \nati{(see Figure 9 of the manuscript)}.
It can probably be explained by the fact that 2DCNN breaks the rotational invariance of the data. 
The problem comes from the fact that the projection introduce a distortion in the geometry and the network has to learn a correct weighting scheme to compensate for it. 
Unfortunately, this operation waste many parameters, reducing the modelization power of the network and requires large amount of training data which is not fully available.
Note that, when visualizing the learned convolution kernels, we observed that they were not radial.

% \TK{makes sense to me.}

% \todo{experiment in prograss by nati}

% \nati{\textbf{Very important point!!!}
% \begin{itemize}
%   \item We can do this comparison...
%   \item Since we used 1/12 of the sphere, all the pixel grids we use are somehow squared. I am now using this as a projection. @Tomek, do you know if this is ok? Or do you know how to do the projection in a simple and efficient manner?
%   \item I think that traditional 2D CNNs will be as good as our algorithm. We should be very careful in the way we formulate this in our answer and in the paper.
% \end{itemize}
% }
% \TK{I would be tempted to leave it to future work! It is an interesting comparison, but requires more work and it brings only a bit more to already awesome paper.}

% \mdeff{Agreed that we should do at least one thing they asked for.
% \begin{itemize}
% 	\item Testing on the 1/12 of the sphere is not what he asked for. On one hand, a global projection would exhibit more distortion. On the other, we never use the whole sphere.
% 	\item Keep in mind that there is no orientation that is consistent for the 12 base faces (i.e., where is up?).
% 	\item Agreed that the performance should be very similar. Both are doing local computations. And locally (3 to 5 pixels across), the sphere is isomorphic to an Euclidean plane.
% 	\item The only difference, in my opinion, would be due to the anisotropic filters the 2D CNN has access to.
% 	\item Not sure we should include this result in the main paper. Hard to justify in the story, especially as a comparison with alternative spherical CNNs is missing. Let's see the results.
% \end{itemize}
% }

@@

- The first "is" is superfluous.
\end{mdframed}
This has been corrected.

\subsection{}

\begin{mdframed}[style=comment]

Finally, after all these little points, I just want to congratulate the authors again on a great paper. It was a pleasure reading it. I will be sure to be using this method in the near future!
\end{mdframed}
We would like to thank the reviewer again for his exceptionally detailed review. We really appreciate his effort and the time he invested in his answers. 
He is welcome to contact us for additional information or discussion when the publication process will be over.

### 2019-03-04 scnn.tex
comment 2.2

@@

\figref{results} compares the classification accuracy of DeepSphere (FCN and CNN variants) and the two baselines across five noise levels and three sample sizes.
We indeed observe that the problem is made more difficult as the sample size decreases, and the noise increases.
As fewer pixels are available to make a decision about a sample, the algorithms have access to less information and thus cannot classify as well.
As the noise increases, the useful information gets diluted in irrelevant data, i.e., the signal-to-noise ratio (SNR) diminishes.

@@

@@

>>>>

\nati{
Additionally, we compare DeepSphere to a traditional 2-dimensional convolutional neural network (2DCNN). 
To do so, the samples are first transformed into a square image before being fed into the 2DCNN. 
As described in \ref{sec:projection}, we use the geometrical properties of HEALPix to construct an efficient and simple projection of the samples. 
Note that during the review process another spherical CNN architecture using exploiting this property appeared \cite{krachmalnicoff2019convolutional}. 
We used convolution kernel of size $5\times 5$ and reduced the number of filters in order to match the number of variables between DeepSphere and the 2DCNN.
}
\nati{Say something about the number of parameters/fitlers. }

@@

\figref{results} compares the classification accuracy of DeepSphere (FCN and CNN variants), the 2DCNN and the two baselines across five noise levels and three sample sizes.
We indeed observe that the problem is made more difficult as the sample size decreases, and the noise increases.
As fewer pixels are available to make a decision about a sample, the algorithms have access to less information and thus cannot classify as well.
As the noise increases, the useful information gets diluted in irrelevant data, i.e., the signal-to-noise ratio (SNR) diminishes.

@@

\nati{
To our surprise, the traditional 2DCNN did not match the performance of DeepSphere. 
It can probably be explained by the fact that 2DCNN breaks the rotational invariance of the data. 
The problem comes from the fact that the projection introduce a distortion in the geometry and the network has to learn a correct weighting scheme to compensate for it. 
Unfortunately, this operation waste many parameters, reducing the modelization power of the network and requires large amount of training data which is not fully available.
Note that, when visualizing the learned convolution kernels, we observed that they were not radial. 
}

@@

\nati{
\section{Maps projection onto a 2-dimensional plane}
\label{sec:projection}
In order to compare the DeepSphere architecture with a traditional 2d-CNN, we need to project the spherical data onto a plane. 
Problematically a radial projection is computationally expensive to the point where, because of the dataset augmentation, it becomes significantly more expensive than the actual training of the network.
Hence, we decided to use the geometrical properties of HEALPix to create a ``special'' projection. 
In the HEALPix sampling, the sphere is first separated into $12$ parts that are then subdivided into $4$ recursively.
Eventually, each of these $12$ part resembles a square grid. 
To visually see it, we refer the reader to \figref{border_effects}.
As the problem we are solving only involve one of these 12 parts, we can use this natural structure to simply re-assign each HEALPix pixel to one of a 2 dimensional grid of size $N_{sides} \times N_{sides}$. 
This operation is computationally very cheap and probably spread quite evenly across the resulting image. 
This projection has recently been used for a CNN in cosmology in the very recent work \cite{krachmalnicoff2019convolutional}.  
}
